---
title: "tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

**MRTpostInfLASSO** is an R package for estimating causal excursion effects in micro-randomized trials and conducting valid post-selection inference after variable selection. This vignette provides a quick overview of the function DR_WCLS_LASSO in the **MRTpostInfLASSO** package. DR-WCLS is a doubly robust inferential procedure for estimating time-varying causal effects in longitudinal data that incorporate debiased machine learning techniques. The algorithm characterizes the problem as a two-stage weighted regression estimation that regresses the estimated pseudo-outcomes on a feature vector. Specifically, it solves

$$
\frac{1}{\sqrt{n}}
\sum_{i=1}^{n}\sum_{t=1}^{T}
\Bigl[
  \hat{p}^{(k)}_{t}(1 \mid S_t)\,
  \bigl(1-\hat{p}^{(k)}_{t}(1 \mid S_t)\bigr)\,
  \bigl(\tilde{Y}^{(\mathrm{DR})}_{i,t+1}-f_t(S_t)^{\top}\beta\bigr)^2
\Bigr]
\;+\;
\lambda \lVert \beta \rVert_{1}
\;-\;
w^{\top}\beta
$$
.


 
## Installation

```{r warning = FALSE}
# Install From GitHub
# install.packages("devtools")
# devtools::install_github("WHD-Lab/DR_WCLS_LASSO")
```

## Loading the Package

```{r}
library(MRTpostInfLASSO)
```

## Example

We will use the `HeartSteps` dataset to illustrate.

```{r}
# Load HeartSteps Data
library(MRTAnalysis)
data(data_mimicHeartSteps)
# head(data_mimicHeartSteps)

```


```{r}
set.seed(100)
Ht = c('logstep_30min_lag1','logstep_pre30min','is_at_home_or_work','day_in_study')
St = c('logstep_30min_lag1','logstep_pre30min','is_at_home_or_work','day_in_study')
At = 'intervention'
outcome = 'logstep_30min'
UI_return = DR_WCLS_LASSO(data = data_mimicHeartSteps, 
                          fold = 5, ID = "userid", 
                          time = "decision_point", Ht = Ht, St = St, At = At, 
                          prob = 'rand_prob', outcome = outcome,
                          virtualenv_path = 'fakepath/wcls',
                          method_pesu = "CVLASSO", lam = NULL,
                          noise_scale = NULL, splitrat = 0.8,
                          level = 0.9)

```

Intern Health Study

```{r}

# df = read.csv('~/Desktop/25Winter/IHS Design/Hyperparameter/LogStepsReward.csv')
# 
# Ht = c('StepsPastDay', 'Steps1HourBefore', 'is_weekend', 'StepsPastWeek','StepsPast3Days', 'maxHRPast24Hours', 'RHR', 'HoursSinceLastMood')
# St = c('StepsPastDay', 'Steps1HourBefore', 'is_weekend', 'StepsPastWeek','StepsPast3Days', 'maxHRPast24Hours', 'RHR', 'HoursSinceLastMood')
# At = 'sent'
# outcome = 'LogStepsReward'
# 
# df$prob = rep(0.5, length(df$PARTICIPANTIDENTIFIER))
# 
# df_mini = df[1:860,]
# UI_return_IHS = DR_WCLS_LASSO(data = df, 
#                           fold = 5, ID = "PARTICIPANTIDENTIFIER", 
#                           time = "time", Ht = Ht, St = St, At = At, 
#                           prob = 'prob', outcome = outcome,
#                           virtualenv_path = 'fakepath/wcls',
#                           method_pesu = "CVLASSO", lam = NULL,
#                           noise_scale = NULL, splitrat = 0.8,
#                           level = 0.9, core_num = 3)

```


